\section{Analysis in $\mathbb{R}^{n}$}

Metric / Distance Function, Vector Space, Euclidean Space, Euclidean Norm, Euclidean Distance.

Open Ball, Interior Point, Exterior Point, Boundary Point, Open Set, Closed Set, Closure Point, Accumulation Point, Isolated Point.

Sequence, Sequence Converge, Sequence Bounded, Subsequence.

Compact, Sequentially Compact, \underline{$\text{Compact} \iff \text{Sequentially Compact}$}, 

\underline{$\text{Compact} \iff \text{Closed and Bounded}$}, \underline{\textbf{Bolzano-Weierstrass Theorem}}, Cauchy Sequence, Complete, \underline{$\mathbb{Q}$}.

Limits of a Function (definition, how to prove not exists), Continuity (definition, use properties to prove continuity), \underline{Weierstrass Theorem}.

\section{Multivariate Analysis}

Derivate with Direction, Partial Derivative, Gradient, Directional Derivative, Differentiable, Differential, \underline{Differential, Continuous, Partial Derivates Exist}.

Jacobian Matrix, Chain Rule, Sufficient Condition of Differentiability.

Level Set, Tangent Hyperplane.

Hessian Matrix, \textbf{Taylor Formula in $\mathbb{R}^{n}$}, Inverse Function Theorem (conditions), Implicit Function Theorem (conditions).

Convex Set, Convex Envelope, Separating Hyperplane, Strict / Weak Separation Theorem.

Convex / Concave Function (definition, properties, FOC, SOC), Upper-level / Lower-level Set, Quasi-convex / Quasi-concave Function (definition, examples, properties).


\section{Static Optimization}

\subsection{Unconstrained}
FOC (critical / stationary point, maximizer, minimizer, saddle point)
SOC (using Hessian Matrix's definiteness)
Convex Program (global maximizer and minimizer).

\subsection{Equality Constraints}
Lagrange Method, FOC, SOC, Convex Case.

\subsection{Inequality Constraints}
FOC(KKT Conditions), SOC, Convex Case.

\section{Difference and Differential Equations}

Asymptotically Stable, Unstable, Marginally Stable.

The Solutions of N-order Linear Difference Equations with Constant Coefficients.

First-order System of Difference Equations ($A(k), A$).

The Solutions of 1st-order Differential Equations (Separable, Change of variable, linear).

The Solutions of N-order Linear Differential Equations with Constant Coefficients.

First-order System of Differential Equations ($A(t), A$).


